An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-08-02 EconGrowth2.Rnw
Edited the interpretation text from Martin.
Added a description of the resampling process.

**** To be completed by Randy Prium ****

Discuss resampling theory.

Discuss our resampling approach.

Be sure to include references on resampling. 
Resampling will be new territory for the readers of \emph{Energy Economics}

@@

\begin{itemize}
\item{Use CD as starting points and do CES convergence on that basis.}
\item{Do scatterplots to get range of uncertainty inputs, then re-interpret. delta against sigma plots, as well as $\delta_1$ and $\sigma_1$ plots.}
\end{itemize}

@@

CES is a broader specification of CD, LINEX is not. In LINEX no delta's nor sigmas are generated. Only efficiency terms are generated in LINEX, thus in other words economic growth is only possible by increased efficiency. 
The efficiency of using energy is driven by capital stock. The objective/growth engine is to increase capital at a rate that is faster then growth in labour and energy.  Increasing labour relative to energy gives a muted benefit. 
LINEX derived for Germany: only really applicable when energy and capital are complements.

>>>>

Resampling is a statistical technique for estimating the 
precision of coefficient estimates. We implemented
residual resampling in our parameter estimation processes.

When resampling, the modeler assumes that the residuals ($r_t$) between 
observed data (in our case historical GDP values, $y_{act,t}$) and
model predictions (in our case, $y_{pred,t}$) at
time $t$ are representative 
of the uncertainty and random scatter in the observed historical data.

\begin{equation} \label{eq:residual}
  r_t = y_{pred,t} - y_{act,t}
\end{equation}

Under that assumption, it would not be surprising if the residual from 
any one year had been observed on a different year. 
By applying randomly selected residuals (with randomly selected signs)
to model predictions at each year of the GDP time series, we create a resampled
GDP time series ($y_{resample,t}$) that can be re-fitted using the techniques
described below.

\begin{equation} \label{eq:resample_y}
  y_{resample,t} = y_{pred,t} + \mathrm{rand}(-1, 1) \cdot \mathrm{abs}(\mathrm{rand}(r_t))
\end{equation}

We repeated the resampling process \Sexpr{numResamples()} times for each
combination of growth model, energy quantification, and country.

The coefficients from the fit to the resampled time series
(the ``resample coefficients'') will be different from the coefficients 
obtained from the fit to historical data (the ``base coefficients'').
Large residuals in the base fit
are expected to produce large variation in resample coefficients, 
indicating less certainty in the knowledge
of those coefficients. 
Similarly, small base fit residuals are expected to produce small
variation in resample coefficients, indicating better knowledge
of the model coefficients.

It is important to note that large residuals can arise from 
(a) poor quality historical time series data ($y_{act}$) or 
(b) a mathematical model that does not describe the underlying phenomena well.
In either case, we would not want to claim precise knowledge of the 
fitted coefficients for a model. The resampling technique that we employed
provides looser coefficient distributions when either of the above
is true. After discussing the fitting technique for each 
growth model, we show scatter plots of fitted coefficients
to provide visual indication of the uncertainty 
of the coefficients for each model.

**** Randy Pruim review the above text and supply some text and references. ****

@@

\begin{itemize}
\item{Use CD as starting points and do CES convergence on that basis.}
\item{Do scatterplots to get range of uncertainty inputs, then re-interpret. 
$\delta$ against $\sigma$ plots, as well as $\delta_1$ and $\sigma_1$ plots.}
\end{itemize}

@@

CES is a broader specification of CD, LINEX is not. 
In LINEX neither $\delta$s nor $\sigma$s are generated. 
Only efficiency terms are generated in LINEX, 
thus in other words economic growth is only possible by increased efficiency. 
The efficiency of using energy is driven by capital stock. 
The objective/growth engine is to increase capital 
at a rate that is faster then growth in labour and energy.  
Increasing labour relative to energy gives a muted benefit. 
LINEX derived for Germany: only really applicable when energy and capital are complements.
